### 2020-10-26 PANNEURO_CFS_merge_script_WHO-pres.Rmd

---
title: "PANCHANGE_data_merging"
author: "Christopher Huebel"
date: "17/08/2020"
output: html_document
---

This is the preprocessing script to prepare the data to run analyses for the project "Anxiety, depression and trauma symptom change during the COVID-19 pandemic: retrospective versus objective assessment" - Young et al (2020)

Script written by K Purves, K Thompson, C Huebel and M Davies.
Email: kirstin.purves@kcl.ac.uk, katie.thompson@kcl.ac.uk, christopher.1.huebel@kcl.ac.uk, molly.davies@kcl.ac.uk

#Set up
```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      comment=NA,
                      prompt=FALSE,
                      cache=FALSE)

options(bitmapType = 'quartz') # to render fonts better
```

Clear global environment
```{r Clear global environment}
remove(list = ls())
```

Retrieve the current date to use it for file endings to not overwrite files when one exports files
```{r Recent date}
date = Sys.Date()
```

Load packages
```{r}
library(tidyverse)
library(skimr)
```

# Source data file paths

```{r source file path}
#source raw data directory: data.raw_path
source("../PANCHANGE_raw_path.R")
```

# Import data
```{r import data}
coping_baseline <- read_rds(path = paste0(data_path, "/four_cohorts.rds"))
dim(coping_baseline)

cfs_wave6 <- read_rds(path = paste0(data_path, "/cfs_wave6.rds"))
dim(cfs_wave6)

```

```{r merge data frames}
dat <- left_join(
  coping_baseline,
  cfs_wave6, by = c("Sample", "ID")
)
```

```{r Skim data}
#look at the data 
#skimr::skim(dat) #(too big to run)
dim(dat)
```

```{r Save four cohorts as rds file}
saveRDS(object = dat, file = paste0(data_path, "/baseline_and_cfs_wave6.rds"))
```