### 2020-10-27 PANNEURO_preprocessing_BMI_WHO-pres.Rmd

---
title: "PANNEURO_CFS_cleaning_WHO_pres"
author: "Molly Davies (adapting from a script by K Purves, K Thompson, and C Huebel)"
date: "26/10/2020"
output: html_document
---

This is a script to quickly clean the Chalder Fatigue Scale in COPING follow-up set B, for the purpose of Gerome Breen's presentation to the WHO.

#Set up
```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      comment=NA,
                      prompt=FALSE,
                      cache=FALSE)

options(bitmapType = 'quartz') # to render fonts better
```

Clear global environment
```{r Clear global environment}
remove(list = ls())
```

Retrieve the current date to use it for file endings to not overwrite files when one exports files
```{r Recent date}
date = Sys.Date()
```

#Packages
Install packages (if they are not available in your version of R)
```{r Installing packages}
#install.packages("summarytools")
#install.packages("tidyverse")
#install.packages("psych")
#install.packages("eeptools")
#install.packages("skimr")
```

Load packages
```{r Load packages}
library(knitr)
library(summarytools)
library(psych)
library(eeptools) # Needed to calculate age from birthdate
library(tidyverse)
```

# Colour palettes
Define colours for plotting this are the standard coping colours
```{r Colour palettes: COPING}
COPINGpalette2 <- c("#78D9C5",
                    "#F5BE5E")

COPINGpalette3 <- c("#78D9C5",
                    "#F5BE5E",
                    "#EEB6E9")

COPINGpalette4 <- c("#78D9C5",
                    "#F5BE5E",
                    "#EEB6E9",
                    "#DBDB73")

COPINGpalette5 <- c("#78D9C5",
                    "#F5BE5E",
                    "#EEB6E9",
                    "#DBDB73", 
                    "#FFED98")

COPINGpalette6 <- c("#78D9C5",
                    "#F5BE5E",
                    "#EEB6E9",
                    "#DBDB73",
                    "#FFED98",
                    "#BFD2EB")

COPINGpalette7 <- c("#78D9C5",
                    "#F5BE5E",
                    "#EEB6E9",
                    "#DBDB73", 
                    "#FFED98",
                    "#BFD2EB", 
                    "#808080")

COPINGpaletteGRAD <- c("#F5BE5E",
                       "#FFD284",
                       "#FFEED1",
                       "#B5B5B5",
                       "#DEFFF8",
                       "#94F6E1",
                       "#78D9C5")

COPINGNeuCenterpalette <- c("#78D9C5",
                            "#808080",
                            "#F5BE5E")

RAMPworseGRADpalette <- c("#78D9C5",
                          "#FFEED1",
                          "#F5BE5E",
                          "#FFB1B5")
GLADpalette = c("#efc00b", 
                "#b7dee8")  
```

Choose in this chunk which palette to use
04.07.2020 - Default to 2 colour COPING palette
```{r Choose colour palette}
palette = COPINGpalette2
```

#Add_numeric function
Function used to convert character variables into numeric variables. 
```{r add_numeric function}
add_numeric <- function(dat, exclude = NULL) {
  dat_fct <- sjlabelled::as_label(dat)

  dat <- dat[!colnames(dat) %in% exclude]
  colnames(dat) <- paste(colnames(dat), "numeric", sep = "_")
  return(bind_cols(dat_fct, dat))
}
```

# Data import and merging

Note: GLAD only receives current measures and RAMP questionnaire
Note: All COPING surveys are separate surveys

# Source data file paths

```{r source file path}
#source raw data directory: data.raw_path
source("../PANCHANGE_raw_path.R")
```

# Read in data
```{r glad dem cols}
nbr.bmi.raw <- readRDS(file = paste0(data.raw_path, "/nbr/dem_nbr.rds"))
dim(nbr.bmi.raw)
colnames(nbr.bmi.raw)

glad.bmi.raw <- readRDS(file = paste0(data.raw_path, "/glad/dem_glad.rds"))
dim(glad.bmi.raw)
colnames(glad.bmi.raw)

edgi.bmi.raw <- readRDS(file = paste0(data.raw_path, "/edgi/dem_edgi.rds"))
dim(edgi.bmi.raw)
colnames(edgi.bmi.raw)
```

# Select variables
```{r cfs cleaning}

# NBR cohort
nbr.bmi.raw <- nbr.bmi.raw %>%
    drop_na(subjectid) %>% # Drop NAs
    distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(subjectid, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
    separate(ID.intm, into = "ID", sep = 7) %>%
    mutate(ID = as.numeric(ID)) %>%
    select(
      Sample, # Sample
      ID, # ID
         height.1 = demographics.what_is_your_current_height.1, 
         height.2 = demographics.what_is_your_current_height.2, 
         height = demographics.what_is_your_current_height, 
         weight = demographics.provide_current_weight_pregnant,
         weight.1 = demographics.provide_current_weight_pregnant.1, 
         weight.2 = demographics.provide_current_weight_pregnant.2)

# GLAD cohort
glad.bmi.raw <- glad.bmi.raw %>%
    drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID)) %>%
    select(
      Sample, # Sample
      ID, # ID
         height.1 = demographics.what_is_your_current_height.1, 
         height.2 = demographics.what_is_your_current_height.2, 
         height = demographics.what_is_your_current_height, 
         weight = demographics.current_weight_pregnant_weight,
         weight.1 = demographics.current_weight_pregnant_weight.1, 
         weight.2 = demographics.current_weight_pregnant_weight.2
         )

# EDGI

edgi.bmi.raw <- edgi.bmi.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
    distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
    mutate(ID = as.numeric(ID)) %>%
    select(
      Sample, # Sample
      ID, # ID
         height.1 = demographics.what_is_your_current_height.1, 
         height.2 = demographics.what_is_your_current_height.2, 
         height = demographics.what_is_your_current_height, 
         weight = demographics.pregnant_current_weight_provide,
         weight.1 = demographics.pregnant_current_weight_provide.1, 
         weight.2 = demographics.pregnant_current_weight_provide.2
         )

 colnames(edgi.bmi.raw) <- c("Sample", "ID", "inch", "cm", "ft", "stn", "pnd", "kg")
 colnames(nbr.bmi.raw) <- c("Sample", "ID", "inch", "cm", "ft", "stn", "pnd", "kg")
 colnames(glad.bmi.raw) <- c("Sample", "ID", "ft", "inch", "cm", "kg", "stn", "pnd")

edgi.bmi.raw$ft_cm <- (edgi.bmi.raw$ft * 30.48)
edgi.bmi.raw$inch_cm <- (edgi.bmi.raw$inch * 2.54)
edgi.bmi.raw$new_cm <- (edgi.bmi.raw$ft_cm + edgi.bmi.raw$inch_cm)
edgi.bmi.raw$cm_all <- rowSums(edgi.bmi.raw[, c("cm", "new_cm")], na.rm=T)

edgi.bmi.raw$stn_kg <- (edgi.bmi.raw$stn * 6.35029)
edgi.bmi.raw$pnd_kg <- (edgi.bmi.raw$pnd * 0.453592)
edgi.bmi.raw$new_kg <- (edgi.bmi.raw$stn_kg + edgi.bmi.raw$pnd_kg)
edgi.bmi.raw$kg_all <- rowSums(edgi.bmi.raw[, c("kg", "new_kg")], na.rm=T)

nbr.bmi.raw$cm <- as.numeric(nbr.bmi.raw$cm)
nbr.bmi.raw$ft_cm <- (nbr.bmi.raw$ft * 30.48)
nbr.bmi.raw$inch_cm <- (nbr.bmi.raw$inch * 2.54)
nbr.bmi.raw$new_cm <- (nbr.bmi.raw$ft_cm + nbr.bmi.raw$inch_cm)
nbr.bmi.raw$cm_all <- rowSums(nbr.bmi.raw[, c("cm", "new_cm")], na.rm=T)

nbr.bmi.raw$stn_kg <- (nbr.bmi.raw$stn * 6.35029)
nbr.bmi.raw$pnd_kg <- (nbr.bmi.raw$pnd * 0.453592)
nbr.bmi.raw$new_kg <- (nbr.bmi.raw$stn_kg + nbr.bmi.raw$pnd_kg)
nbr.bmi.raw$kg_all <- rowSums(nbr.bmi.raw[, c("kg", "new_kg")], na.rm=T)

glad.bmi.raw$ft_cm <- (glad.bmi.raw$ft * 30.48)
glad.bmi.raw$inch_cm <- (glad.bmi.raw$inch * 2.54)
glad.bmi.raw$new_cm <- (glad.bmi.raw$ft_cm + glad.bmi.raw$inch_cm)
glad.bmi.raw$cm_all <- rowSums(glad.bmi.raw[, c("cm", "new_cm")], na.rm=T)

glad.bmi.raw$stn_kg <- (glad.bmi.raw$stn * 6.35029)
glad.bmi.raw$pnd_kg <- (glad.bmi.raw$pnd * 0.453592)
glad.bmi.raw$new_kg <- (glad.bmi.raw$stn_kg + glad.bmi.raw$pnd_kg)
glad.bmi.raw$kg_all <- rowSums(glad.bmi.raw[, c("kg", "new_kg")], na.rm=T)

# finally, we can calculate the BMI

edgi.bmi.raw$cm_whole <- edgi.bmi.raw$cm_all/100
edgi.bmi.raw$bmi <- edgi.bmi.raw$kg_all / (edgi.bmi.raw$cm_whole^2)

nbr.bmi.raw$cm_whole <- nbr.bmi.raw$cm_all/100
nbr.bmi.raw$bmi <- nbr.bmi.raw$kg_all / (nbr.bmi.raw$cm_whole^2)
nbr.bmi.raw$bmi_new <- round(nbr.bmi.raw$bmi, digits = 2)
nbr.bmi.raw$bmi <- nbr.bmi.raw$bmi_new

glad.bmi.raw$cm_whole <- glad.bmi.raw$cm_all/100
glad.bmi.raw$bmi <- glad.bmi.raw$kg_all / (glad.bmi.raw$cm_whole^2)

# then let's get this into the main COPING data frame and delete this one and also tidy up

# Merge datasets

glad.bmi.raw <- glad.bmi.raw %>%
  select(Sample, ID, bmi)

edgi.bmi.raw <- edgi.bmi.raw %>%
  select(Sample, ID, bmi)

nbr.bmi.raw <- nbr.bmi.raw %>%
  select(Sample, ID, bmi)

 bmi <- as.data.frame(rbind(nbr.bmi.raw, 
                                 glad.bmi.raw, 
                                 edgi.bmi.raw))

#Export dataset

saveRDS(object = bmi, file = paste0(data_path, "/bmi.rds"))