### 2020-12-04 Retro and prepan comparisons.Rmd

---
title: "Retro and prepan comparisons"
author: "Steven Bright and Henry Rogers"
date: "30/10/2020"
output: html_document
---

#Set up

```{r clean environment}

remove(list = ls())

```

```{r Save a date object}

date <- Sys.Date()

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Set WD

```{r Set WD}

#Steven's working environment
setwd("c:\\Users\\Student\\Documents\\Git\\PANCHANGE_analysis")
getwd()

#Henry's working environment

```

#Load packages

```{r Load packages}

library(tidyverse)
library(lfactors)
library(data.table)

```

#Set file paths and data names

```{r Source data and file paths}
#Sourcing data paths for raw data (data_path) and clean data (data_path_clean)
#Sourcing data files for 1) saliva_kits 2) coping_glad 3) demographic datasets

source(file = "../Coping_data_path.R")

#May need a different path for PC computers
#source(file = "..\\flare_data_path.R")

```

#Prepare data for merge

```{r Read in data}

data_glad <- read_csv(paste0(data_path, coping_glad))
data_nbr <- read_csv(paste0(data_path, coping_nbr))
data_edgi <- read_csv(paste0(data_path, coping_edgi))

```

```{r Drop first 2 rows}

coping_glad_cohort <- data_glad[3:dim(data_glad)[1],]
coping_edgi_cohort <- data_edgi[3:dim(data_edgi)[1],]
coping_nbr_cohort <- data_nbr[3:dim(data_nbr)[1],]

```

Perform the left join

```{r left join}

coping_data <- left_join(coping_nbr_cohort, 
                         coping_glad_cohort,
                         by = "ExternalReference")

coping_data <- left_join(coping_data, 
                         coping_edgi_cohort,
                         by = "ExternalReference")

#MD addition to visualise:
head(coping_data)

```

#Task overview

1) No change retrospective scores: "impute" to be the same as their pandemic baseline

2) Check correlation between these and the pre-pandemic scores in COPING

3) ACCURACY SCORES: Another question: what proportion of the time are people right about

worse / better / no change (compared to pre-pandemic)

Two categories: (1) is it accurate (i.e. if they say better are they at all better) and (2) is it more than 3 points different

#Task 1: Impute no change scores