### 2020-12-04 PANCHANGE_data_merging.Rmd
small tweak during review. New knit

---
title: "PANCHANGE_data_merging"
author: "Christopher Huebel"
date: "17/08/2020"
output: html_document
---

This is the preprocessing script to prepare the data to run analyses for the project "Anxiety, depression and trauma symptom change during the COVID-19 pandemic: retrospective versus objective assessment" - Young et al (2020)

Script written by K Purves, K Thompson, C Huebel and M Davies.
Email: kirstin.purves@kcl.ac.uk, katie.thompson@kcl.ac.uk, christopher.1.huebel@kcl.ac.uk, molly.davies@kcl.ac.uk

#Set up
```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      comment=NA,
                      prompt=FALSE,
                      cache=FALSE)

options(bitmapType = 'quartz') # to render fonts better
```

Clear global environment
```{r Clear global environment}
remove(list = ls())
```

Retrieve the current date to use it for file endings to not overwrite files when one exports files
```{r Recent date}
date = Sys.Date()
```

Load packages
```{r}
library(tidyverse)
library(skimr)
```

# Source data file paths

```{r source file path}
#source raw data directory: data.raw_path
source("../PANCHANGE_raw_path.R")
```

# Import data
```{r export data}
glad <- read_rds(paste0(data_path, "glad.rds"))
dim(glad)

edgi <- read_rds(paste0(data_path, "edgi.rds"))
dim(edgi)

nbr <- read_rds(paste0(data_path, "nbr.rds"))
dim(nbr)

ramp <- read_rds(paste0(data_path, "ramp.rds"))
dim(ramp)
```

glad/edgi data frame list
```{r glad/edgi data frame list}
dat <- bind_rows(
  glad,
  edgi,
  nbr,
  ramp
)
```

```{r Skim data}
#look at the data
skimr::skim(dat)
```

```{r Save four cohorts as rds file}
saveRDS(object = dat, file = paste0(data_path, "/four_cohorts.rds"))
```